### 2020-04-07 manuscript/manuscript.RmD
- updated introduction

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    According to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, most findings are based on self-reports, and little is known about whether the privacy calculus can be used to explain observations of actual authentic behavior. Likewise, we still know comparatively little whether the privacy calculus can be influenced by the design of online websites, including for example popularity cues such as like and dislike buttons. To answer these questions, we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites where they could discuss a current political topic. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. The results showed that privacy calculus variables predicted a considerable share of actual self-disclosure. The impact of popularity cues was negligible. In conclusion, the results demonstrate that self-disclosure online can be explained by privacy concerns and specific gratifications. This finding has several implications. For example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

Somewhat surprisingly, despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018] the privacy paradox has garnered little empirical support. A recent meta-analysis revealed that the correlation between privacy concerns and self-disclosure on SNS is *r* = -.13 [@baruhOnlinePrivacyConcerns2017], indicating that privacy concerns are indeed related to self-disclosure online.

Rather than further pursuing the privacy paradox, a large share of current day research posits that self-disclosure online can be explained---at least partly---by means of the so-called _privacy-calculus_ [@krasnovaOnlineSocialNetworks2010]. The privacy calculus builds on the work of @lauferPrivacyConceptSocial1977 and claims that both expected risks *and* expected benefits explain self-disclosure. Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information [@heirmanPredictingAdolescentsDisclosure2013; koohikamaliInvestigationDynamicModel2019].

However, although the privacy calculus has gained some momentum several important questions remain unanswered. 
First, we still know comparatively little about whether the privacy calculus can be replicated with actual behavioral data in an authentic long-term setting [@kokolakisPrivacyAttitudesPrivacy2017]. 
Thus far, most research supporting the privacy calculus has used either self-reports of behavior [e.g., @krasnovaOnlineSocialNetworks2010], vignette approaches [e.g., @bolUnderstandingEffectsPersonalization2018], or one-shot experiments in the lab [e.g., @treptePrivacyCalculusContextualized2020]. 
However, all three of these approaches significantly hamper external validity.

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for any substantial or explicit weighing of pros and cons.

We agree and consider it necessary to now explicitly focus on the privacy deliberation process itself. 
Moreover, and on a more general level, we aim to gauge the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-disclosure self-efficacy.

Finally, we want to determine whether the privacy calculus can be affected by the design of a website. Specifically, we analyze whether *popularity cues* such as like and dislike buttons affect self-disclosure and the privacy calculus.

>>>>

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    According to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, most findings were based on self-reports, and little is known about whether the privacy calculus can explain also observations of authentic behavior. Likewise, we still know comparatively little as to whether the privacy calculus can be influenced by the design of online websites, including for example popularity cues such as like and dislike buttons. To answer these questions, we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites, on which they could discuss a current political topic. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. Although the originally preregistered model could not be confirmed, the results showed that the privacy calculus variables predicted a considerable share of actual self-disclosure. The impact of popularity cues was negligible. In conclusion, the results indicate that self-disclosure online can be explained by privacy concerns and specific gratifications. This finding has several implications. For example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, field experiment, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

Somewhat surprisingly, despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018] the privacy paradox has garnered little empirical support. A recent meta-analysis revealed that the correlation between privacy concerns and self-disclosure on SNS is *r* = -.13 [@baruhOnlinePrivacyConcerns2017], which indicates that privacy concerns are indeed related to self-disclosure online.

Rather than further pursuing the privacy paradox, a large share of current day research hence posits that self-disclosure online can be explained---at least partly---by means of the so-called _privacy-calculus_ [@krasnovaOnlineSocialNetworks2010]. 
Building on the work of @lauferPrivacyConceptSocial1977, the privacy calculus  claims that both expected risks *and* expected benefits explain self-disclosure. Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information [@heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

However, although the privacy calculus has gained some momentum in academic research several important questions remain unanswered. 
First, we still know comparatively little about whether the privacy calculus can be replicated with actual behavioral data in an authentic long-term setting [@kokolakisPrivacyAttitudesPrivacy2017]. 
Thus far, most research supporting the privacy calculus has used either self-reports of behavior [e.g., @krasnovaOnlineSocialNetworks2010], vignette approaches [e.g., @bolUnderstandingEffectsPersonalization2018], or one-shot experiments in the lab [e.g., @treptePrivacyCalculusContextualized2020]. 
However, all three of these approaches have reduced external validity.
As a result, in this study we analyze actual information sharing behavior in an authentic online setting.

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for any substantial or explicit weighing of pros and cons. 
We agree and in this study hence explicitly focus on the privacy deliberation process itself. 
Moreover, and on a more general level, we aim to gauge the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-disclosure self-efficacy.

Finally, because the privacy calculus does not take place in a vacuum, and because it is often argued that self-disclosure can be easilty triggered by external circumstances, we analyze whether the privacy calculus can be affected by the design of a website. 
Specifically, in this study we investigate whether *popularity cues* such as like and dislike buttons have the power to affect the privacy calculus and to foster self-disclosure.
